Review: Adaptive Introgression Expanded the Genetic Base of Crops during post-Domestication Spread

### 2017-07-05 paper.tex
Update paper.tex

The story of Asian rice (\emph{Oryza sativa}) domestication is still debated.
A recent investigation pointed to a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter), with later divergence of the two prominent varietal groups, \emph{japonica} and \emph{indica} \cite{molina2011molecular}.
\cite{vaughan2008evolving} support this view, noting that present-day patterns of rice variation could be explained by a single domestication event in a population with high standing variation, followed by dispersal into sympatry with locally-adapted wild relatives in diverse environments, genetic admixture, and selection for adaptive alleles.
On the other hand, both genetic and archaeobotanical evidence point toward independent domestications of \emph{japonica} and \emph{indica} in the Yangtze Basin and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
An alternative explanation, as described by \cite{Huang2012}, suggests an initial domestication of \emph{japonica} in China followed by formation of the \emph{indica} group through hybridization of \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
Adaptive introgression may in fact be the signal interpreted as independent domestications.

The high genetic diversity within domesticated rice is likely due to introgression from wild relatives both within the domestication center(s) and in new environments where rice has dispersed \cite{second1982origin}, and gene flow between domesticated Asian rice and its wild relatives outside of the Yangtze Basin could have imparted locally adaptive traits \cite{vaughan2008evolving}.
The wild relatives \emph{rufipogon} and \emph{nivara} both maintain high phenotypic diversity and exhibit locally-adaptive traits.
\emph{Rufipogon} reproduction can be either primarily sexual or vegetative, and whereas \emph{rufipogon} is adapted to forested wetland environments, \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure.
Likewise, domesticated rice varieties display patterns of local adaptation.
For examples, two of the domesticated rice deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river, the \emph{japonicas} are split into temperate and tropical subgroups, and the \emph{indicas} are best suited for lowland environments.
Gene flow between these wild and cultivated rices, though asymmetrical, is exceedingly common, producing nuisance weedy hybrids when the two are grown in sympatry.

>>>>

The story of Asian rice (\emph{Oryza sativa}) domestication is still debated.
Interspecific hybridization between wild and domesticated rices both ancient and recent has produced a phylogeny that defies clear and simple models.
On one hand, some genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups, \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
Others support a single center of domestication, arguing that present-day patterns of rice variation could be explained by a single domestication event in a population with high standing variation, followed by dispersal into sympatry with locally-adapted wild relatives in diverse environments, genetic admixture, and selection for adaptive alleles \cite{vaughan2008evolving}.
A recent investigation utilizing SNP data from wild and domesticated rice accesssions pointed to a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from rufipogon, with later divergence of indica and japonica \cite{molina2011molecular}.
Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure distance between wild and domesticated rices at and around domesication loci for various geographical locations, finding that \emph{japonica} was likely domesticated in the Pearl River area of Guangxi province, China (just south of the Yangtze Basin), and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia. %Also, Huang includes the following quote: "We noticed that many loci with strong signals of selection were nearly identical in both indica and japonica where FST between indica and japonica was extremely low, indicating that introduction of traits during domestication has in many cases involved introgression events."

%The former first paragraph, replaced from above:
%The story of Asian rice (\emph{Oryza sativa}) domestication is still debated.
%A recent investigation pointed to a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter), with later divergence of the two prominent varietal groups, \emph{japonica} and \emph{indica} \cite{molina2011molecular}.
%\cite{vaughan2008evolving} support this view, noting that present-day patterns of rice variation could be explained by a single domestication event in a population with high standing variation, followed by dispersal into sympatry with locally-adapted wild relatives in diverse environments, genetic admixture, and selection for adaptive alleles.
%On the other hand, genetic and archaeobotanical evidence point toward independent domestications of \emph{japonica} and \emph{indica} in the Yangtze Basin and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
%An alternative explanation, as described by \cite{Huang2012}, suggests an initial domestication of \emph{japonica} in China followed by formation of the \emph{indica} group through hybridization of \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
%Adaptive introgression may in fact be the signal interpreted as independent domestications.

The high genetic diversity within domesticated rice is likely due to introgression from wild relatives both within the domestication center(s) and in new environments where rice has dispersed \cite{second1982origin}, and gene flow between domesticated Asian rice and its wild relatives outside of the Yangtze Basin could have imparted locally adaptive traits \cite{vaughan2008evolving}.
The wild relatives \emph{rufipogon} and \emph{nivara} both maintain high genotypic and phenotypic diversity and exhibit locally-adaptive traits (whereas \emph{rufipogon} is adapted to forested wetland environments, \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure).
%From Huang2012, "Based on the SNP data, the sequence diversity (p) of O. rufipogon was estimated at 0.003, which is higher than that in O. sativa—the sequence diversity is 0.0024 for O. sativa, and 0.0016 and 0.0006 for indica and japonica, respectively21".
Likewise, domesticated rice varieties display patterns of local adaptation (for examples, two of the domesticated rice deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river, the \emph{japonicas} are split into temperate and tropical subgroups, and the \emph{indicas} are best suited for lowland environments.
Gene flow between wild and cultivated rices, though asymmetrical, is frequent enough that nuisance weedy hybrids commonly arise amongst cultivated populations and must be managed with difficulty \cite{vaughan2008evolving}.
